---
title: "Predicting Depression Outcomes Using ML"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: inline
---

# ─────────────────────────────────────────────
# Section 1: Import and Initial Setup
# ─────────────────────────────────────────────

## Install packages
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
  dplyr, tidyr, ggplot2, haven, labelled, readr, here,
  gtsummary, sjPlot, jtools, lubridate, visdat, dataMaid, 
  viridis, kableExtra, flextable, officer, purrr
)

```

## Import data (original repository: xxxxxx)
```{r}

dat <- read_dta(here("data", "THP_clean.dta"))
```

# ─────────────────────────────────────────────
# Section 2: Data preprocessing
# ─────────────────────────────────────────────

### Categorize study samples
```{r}

# View initial breakdown
table(dat$treat, dat$dep_sample)

# Define study group classification
dat <- dat %>%
  mutate(
    study_group = case_when(
      dep_sample == 0 ~ "7-year control",
      (dep_sample == 1 | is.na(dep_sample)) & treat == 1 ~ "Treatment group",
      (dep_sample == 1 | is.na(dep_sample)) & treat == 0 ~ "Baseline control",
      TRUE ~ NA_character_
    )
  )

# Review classification
table(dat$study_group, useNA = "ifany")
```

### Define analysis subsets
```{r}

# Trial participants at baseline (N = 903)
dat_trial_baseline <- dat %>%
  filter(attrit2 %in% c(0, 1))

# At 1-year follow-up (N = 704)
dat_trial_1y <- dat %>%
  filter(attrit %in% c(0, 1))

# At 7-year follow-up (N = 585)
dat_trial_7y <- dat %>%
  filter(attrit2 == 0 & attrit == 0)

# 7-year follow-up full sample: trial + control (N = 885)
dat_7y <- dat %>%
  filter(attrit2 == 0 | dep_sample == 0)

# Control group only (N = 300)
dat_control <- dat %>%
  filter(dep_sample == 0)

table(dat_trial_baseline$study_group)
table(dat_trial_1y$study_group)
table(dat_trial_7y$study_group)
table(dat$study_group)
```


### Define outcomes - cut-off 6 for HAM-D (sensitivity analysis)
```{r, eval= FALSE}

# Percent change and outcomes based on HAM-D ≤ 6 definition
dat_trial <- dat_trial_baseline %>%
  mutate(
    hamd_6m_percentchange = (hamd_6m - hamd_baseline) / hamd_baseline,
    hamd_1y_percentchange = (hamd_1y - hamd_baseline) / hamd_baseline,
    hamd_1y_percentchange_prev = (hamd_1y - hamd_6m) / hamd_6m
  ) %>%
  mutate(across(
    c(hamd_6m_percentchange, hamd_1y_percentchange, hamd_1y_percentchange_prev),
    ~ case_when(
      is.infinite(.) ~ NA_real_,
      near(., 1) ~ 0.999,
      near(., -1) ~ -0.999,
      TRUE ~ .
    )
  )) %>%
  mutate(
    response_6mo = as.integer(hamd_6m_percentchange <= -0.50),
    response_1y = as.integer(hamd_1y_percentchange <= -0.50),
    response_7y = as.integer(depressed == 0),
    
    remission_6mo = as.integer(depressed_6m == 0 & hamd_6m <= 6),
    remission_1y = as.integer(depressed_1y == 0 & hamd_1y <= 6),
    remission_7y = as.integer(depressed == 0),

    recovery_1y = as.integer(depressed_6m == 0 & depressed_1y == 0),
    recovery_7y = as.integer(depressed == 0),

    relapse_1y = as.numeric(case_when(
      remission_6mo == 1 & (depressed_1y == 1 | hamd_1y_percentchange_prev >= 0.5 | hamd_1y > 6) ~ 1L,
      remission_6mo == 1 ~ 0L,
      TRUE ~ NA_integer_
    )),
    
    relapse_7y = as.numeric(case_when(
      remission_1y == 1 & depressed == 1 ~ 1L,
      remission_1y == 1 ~ 0L,
      TRUE ~ NA_integer_
    )),

    recurrence_7y = as.numeric(case_when(
      recovery_1y == 1 & depressed == 1 ~ 1L,
      recovery_1y == 1 ~ 0L,
      TRUE ~ NA_integer_
    ))
  )
```

### Outcomes - cut-off 7 for HAM-D
```{r}

# Adjustment to handle zeros and cutoff 7
dat_trial <- dat_trial_baseline %>%
  mutate(across(
    c(hamd_baseline, hamd_6m, hamd_1y),
    ~ ifelse(. == 0, 0.001, .),
    .names = "{.col}2"
  )) %>%
  mutate(
    hamd_6m_percentchange = (hamd_6m2 - hamd_baseline2) / hamd_baseline2,
    hamd_1y_percentchange = (hamd_1y2 - hamd_baseline2) / hamd_baseline2,
    hamd_1y_percentchange_prev = (hamd_1y2 - hamd_6m2) / hamd_6m2
  ) %>%
  mutate(across(
    c(hamd_6m_percentchange, hamd_1y_percentchange, hamd_1y_percentchange_prev),
    ~ case_when(
      is.infinite(.) ~ NA_real_,
      near(., 1) ~ 0.999,
      near(., -1) ~ -0.999,
      near(., 0) ~ 0.001,
      TRUE ~ .
    )
  )) %>%
  mutate(
    response_6mo = as.integer(hamd_6m_percentchange <= -0.50),
    response_1y = as.integer(hamd_1y_percentchange <= -0.50),
    response_7y = as.integer(depressed == 0),

    remission_6mo = as.integer(depressed_6m == 0 & hamd_6m2 <= 7),
    remission_1y = as.integer(depressed_1y == 0 & hamd_1y2 <= 7),
    remission_7y = as.integer(depressed == 0),

    recovery_1y = as.integer(depressed_6m == 0 & depressed_1y == 0),
    recovery_7y = as.integer(depressed == 0),

    relapse_1y = as.numeric(case_when(
      remission_6mo == 1 & (depressed_1y == 1 | hamd_1y_percentchange_prev >= 0.5 | hamd_1y2 > 7) ~ 1L,
      remission_6mo == 1 ~ 0L,
      TRUE ~ NA_integer_
    )),

    relapse_7y = as.numeric(case_when(
      remission_1y == 1 & depressed == 1 ~ 1L,
      remission_1y == 1 ~ 0L,
      TRUE ~ NA_integer_
    )),

    recurrence_7y = as.numeric(case_when(
      recovery_1y == 1 & depressed == 1 ~ 1L,
      recovery_1y == 1 ~ 0L,
      TRUE ~ NA_integer_
    ))
  )

```


### Rename and label variables
```{r}
variable.names(dat_trial)

########### Set var names
new_names <- c(
  "age_7y" = "age",
  "age_baseline" = "age_baseline",
  "mo_ht_baseline" = "mo_ht",
  "mo_wt_6m" = "mo_wt_6m",
  "mo_bmi_baseline" = "mo_bmi",
  "hamd_baseline" = "hamd_baseline",
  "hamd_baseline2" = "hamd_baseline2",
  "hamd_6m" = "hamd_6m",
  "hamd_6m2" = "hamd_6m2",
  "hamd_1y" = "hamd_1y",
  "hamd_1y2" = "hamd_1y2",
  "depressed_7y" = "depressed",
  "hamd_6m_percentchange_6m" = "hamd_6m_percentchange",
  "hamd_1y_percentchange_1y" = "hamd_1y_percentchange",
  "hamd_1y_percentchange_prev_1y" = "hamd_1y_percentchange_prev",
  "gaf_6m" = "gaf_6m_2",
  "gaf_1y" = "gaf_1y_2",
  "gad_7y" = "gad",
  "location_delivery_7y" = "var619",
  "birth_spacing_7y" = "var621",
  "genhealth_7y" = "genhealth",
  "notunwell_7y" = "notunwell",
  "days_healthy_7y" = "days_healthy",
  "no_hospitalization_7y" = "mo_491_2",
  "no_severe_illness_7y" = "mo_490_2",
  "no_eyesight_problems_7y" = "mo_485_2",
  "no_hearing_problems_7y" = "mo_488_2",
  "ses_baseline" = "ses_bl_flipped",
  "income_hh_7y" = "income_hh",
  "ln_income_fa_7y" = "ln_income_fa",
  "ses_7y" = "ses",
  "edu_mo_7y" = "edu_mo",
  "edu_fa_7y" = "edu_fa",
  "electricity_baseline" = "electricity",
  "radio_baseline" = "radio",
  "tv_baseline" = "tv",
  "fridge_baseline" = "fridge",
  "bicycle_baseline" = "bicycle",
  "motorcycle_baseline" = "motorcycle",
  "aircon_baseline" = "aircon",
  "waterpump_baseline" = "waterpump",
  "cartruck_baseline" = "cartruck",
  "pipedwater_baseline" = "pipedwater",
  "wellwithpump_baseline" = "wellwithpump",
  "wellwater_baseline" = "wellwater",
  "flushtoilet_baseline" = "flushtoilet",
  "anylatrine_baseline" = "anylatrine",
  "brickwalls_baseline" = "brickwalls",
  "enoughfoodmoney_baseline" = "enoughfoodmoney",
  "metalroof_baseline" = "metalroof",
  "bestroof_baseline" = "bestroof",
  "basicneeds_baseline" = "basicneeds",
  "enoughfood_baseline" = "enoughfood",
  "employed_mo_7y" = "employed_mo",
  "empowered_7y" = "empowered",
  "motherfinancial_baseline" = "motherfinancial",
  "motherfinancial_f_7y" = "motherfinancial_f",
  "income_mo_7y" = "income_mo",
  "mspss_tot_7y" = "mspss_tot",
  "grandmother_7y" = "grandmother",
  "maritalscale_7y" = "maritalscale",
  "relationshipindex_7y" = "relationshipindex",
  "nonviolent_7y" = "nonviolent",
  "relationship_husb_7y" = "relationship_husb",
  "relationship_inlaw_7y" = "relationship_inlaw",
  "c_first_child_baseline" = "c_first_child",
  "girl_baseline" = "girl",
  "femshare_baseline" = "femshare",
  "not_stunted_7y" = "not_stunted",
  "c_ht_7y" = "c_ht",
  "c_wt_7y" = "c_wt",
  "WAZ_1y" = "WAZ_12",
  "HAZ_1y" = "HAZ_12",
  "infantdev_7y" = "infantdev",
  "complete_immunization_7y" = "var599",
  "vci_7y" = "vci",
  "vsi_7y" = "vsi",
  "fri_7y" = "fri",
  "wmi_7y" = "wmi",
  "psi_7y" = "psi",
  "fsiq_7y" = "fsiq",
  "sdq_sum_7y" = "sdq_sum",
  "urdu_7y" = "urdu",
  "math_7y" = "math",
  "private_sch_7y" = "private_sch",
  "expect_sch_7y" = "expect_sch",
  "class_size_7y" = "class_size",
  "classroom_7y" = "classroom",
  "home_7y" = "home",
  "stroop_7y" = "stroop",
  "spence_7y" = "spence",
  "motor_7y" = "motor",
  "PPI_7y" = "PPI",
  "home_obs_7y" = "home_obs",
  "play_mo_1y" = "play_mo_1y", 
  "play_fa_1y" = "play_fa_1y"
)

dat_trial_clean <- dat_trial %>%
  dplyr::rename(!!!new_names)

######### Set formats
variable_types <- list(
  ## Factors
  factor = c(
    "treat",
    "newid",
    "uc",
    "interviewer",
    "month_int"
    ),

  ## Numeric
  numeric = c("response_6mo", "response_1y", 
     "remission_6mo",
    "remission_1y", 
    "recovery_1y",
    "recovery_7y",
    "recover_never",
    "relapse_1y",
    "recurrence_7y",
    "hamd_6m_percentchange_6m", 
    "hamd_1y_percentchange_1y", 
    "hamd_1y_percentchange_prev_1y", 
    "depressed_6m", 
    "depressed_1y",
    "depressed_7y",
    "depindex_0", 
    "depindex_0_f", 
    "depindex_6m", 
    "depindex_1y", 
    "depindex_6m_f",
    "depindex_1y_f", 
    "age_baseline",
    "mo_wt_6m",
    "mo_ht_baseline", 
    "mo_bmi_baseline", 
    "hamd_baseline", 
    "hamd_baseline2",
    "hamd_6m", 
    "hamd_6m2",
    "hamd_1y", 
    "hamd_1y2", 
    "bdq_baseline", 
    "bdq_6m", 
    "bdq_1y", 
    "gaf_baseline", 
    "gaf_6m", 
    "gaf_1y",  
    "ses_baseline", 
    "ses_6m", 
    "ses_1y", 
    "employed_fa_baseline", 
    "occupation_fa_baseline", 
    "electricity_baseline", 
    "radio_baseline", 
    "tv_baseline", 
    "fridge_baseline", 
    "bicycle_baseline", 
    "motorcycle_baseline", 
    "aircon_baseline", 
    "waterpump_baseline", 
    "cartruck_baseline", 
    "pipedwater_baseline", 
    "wellwithpump_baseline", 
    "wellwater_baseline", 
    "flushtoilet_baseline", 
    "anylatrine_baseline", 
    "brickwalls_baseline", 
    "enoughfoodmoney_baseline", 
    "metalroof_baseline",
    "basicneeds_baseline", 
    "enoughfood_baseline",  
    "employed_mo_baseline",
    "grandmother_baseline", 
    "grandmother_6m", 
    "grandmother_1y",  
    "famstruct_baseline",
    "nonviolent_1y", 
    "c_first_child_baseline", 
    "girl_baseline",
    "exclusivebf_6m", 
    "anybf_6m",
    "edu_mo_baseline", 
    "edu_fa_baseline",
    "wealth_baseline", 
    "ln_income_fa_6m", 
    "ln_income_fa_1y",
    "motherfinancial_baseline",
    "empowered_6m", 
    "mspss_baseline", 
    "mspss_6m", 
    "mspss_1y", 
    "maritalscale_1y",
    "relationship_husb_1y",  
    "relationship_inlaw_1y", 
    "no_kids_baseline",
    "femshare_baseline",
    "WAZ_6", 
    "WAZ_1y", 
    "HAZ_6", 
    "HAZ_1y", 
    "play_mo_1y", 
    "play_fa_1y")
)


# Convert variables to their appropriate data types and add labels
dat_trial_clean <- dat_trial_clean %>%
  mutate(
    # Convert to factors
    across(.cols = all_of(variable_types$factor), as.factor),
    # Convert to numeric
    across(.cols = all_of(variable_types$numeric), as.numeric))

######### Label variables

# Assign labels to variables
library(labelled)
dat_trial_clean <- dat_trial_clean %>%
  set_variable_labels(
    response_6mo = "Response at 6 months",
    response_1y = "Response at 1 year",
    remission_6mo = "Remission at 6 months",
    remission_1y = "Remission at 1 year",
    recovery_1y = "Recovery at 1 year",
    recovery_7y = "Recovery at 7 years",
    recover_never = "Never recovered",
    relapse_1y = "Relapse at 1 year",
    recurrence_7y = "Recurrence at 7 years",
    treat = "Treatment group", 
    newid = "ID", 
    uc = "Union Council (cluster)", 
    interviewer = "Interviewer", 
    month_int = "Month interviewed", 
    hamd_6m_percentchange_6m = "Percentage change in HAM-D score from baseline at 6 months", 
    hamd_1y_percentchange_1y = "Percentage change in HAM-D score from baseline at 1 year", 
    hamd_1y_percentchange_prev_1y = "Percentage change from 6 months in HAM-D score at 1 year", 
    depindex_0 = "Depression index at baseline", 
    depindex_0_f = "Depression index at baseline (flipped)", 
    depindex_6m = "Depression index at 6 months", 
    depindex_1y = "Depression index at 1 year", 
    depindex_6m_f = "Depression index at 6 months (flipped)", 
    depressed_6m = "Depressed at 6 months", 
    depressed_1y = "Depressed at 1 year", 
    depressed_7y = "Depressed at 7 years",
    depindex_1y_f = "Depressed at 1 year (flipped)", 
    age_baseline = "Mother's age at baseline",
    mo_wt_6m = "Mother's weight at 6 months",
    mo_ht_baseline = "Mother's height at baseline", 
    mo_bmi_baseline = "Mother's BMI at baseline", 
    hamd_baseline = "HAM-D score at baseline", 
    hamd_baseline2 = "HAM-D score at baseline - recoded 0's", 
    hamd_6m = "HAM-D score at 6 months", 
    hamd_6m2 = "HAM-D score at 6 months - recoded 0's", 
    hamd_1y = "HAM-D score at 1 year", 
    hamd_1y2 = "HAM-D score at 1 year - recoded 0's", 
    bdq_baseline = "BDQ at baseline", 
    bdq_6m = "BDQ at 6 months", 
    bdq_1y = "BDQ at 1 year", 
    gaf_baseline = "GAF at baseline", 
    gaf_6m = "GAF at 6 months", 
    gaf_1y = "GAF at 1 year",  
    ses_baseline = "SES at baseline", 
    ses_6m = "SES at 6 months", 
    ses_1y = "SES at 1 year", 
    edu_mo_baseline = "Mother's education at baseline", 
    edu_fa_baseline = "Father's educatino at baseline", 
    employed_fa_baseline = "Father employed at baseline", 
    occupation_fa_baseline = "Father not manual worker at baseline", 
    electricity_baseline = "Household has electricity (baseline)", 
    radio_baseline = "Household has radio (baseline)", 
    tv_baseline = "Household has television (baseline)", 
    fridge_baseline = "Household has fridge (baseline)", 
    bicycle_baseline = "Household has bicycle (baseline)", 
    motorcycle_baseline = "Household has motorcycle (baseline)", 
    aircon_baseline = "Household has airconditioner (baseline)", 
    waterpump_baseline = "Household has water pump (baseline)", 
    cartruck_baseline = "Household has car or truck (baseline)", 
    pipedwater_baseline = "Household has piped water (baseline)", 
    wellwithpump_baseline = "Household has well with pump (baseline)", 
    wellwater_baseline = "Household has well water (baseline)", 
    flushtoilet_baseline = "Household has flush toilet (baseline)", 
    anylatrine_baseline = "Household has any latrine (baseline)", 
    brickwalls_baseline = "Household has brick walls (baseline)", 
    enoughfoodmoney_baseline = "Household has enough food and money (baseline)", 
    metalroof_baseline = "Household has metal roof (baseline)", 
    wealth_baseline = "Household wealth index (baseline)", 
    ln_income_fa_6m = "Father's monthly income (ln) at 6 months", 
    ln_income_fa_1y = "Father's monthly income (ln) at 1 year", 
    basicneeds_baseline = "Have enough for basic needs (baseline)", 
    enoughfood_baseline = "Have enough for food (baseline)",  
    employed_mo_baseline = "Mother usually works (baseline)",
    motherfinancial_baseline = "Mother controls spending (baseline)",
    empowered_6m = "Mother's financial empowerment (6 months)", 
    mspss_baseline = "Perceived social support score (baseline)", 
    mspss_6m = "Perceived social support score (6 months)", 
    mspss_1y = "Perceived social support score (1 year)",  
    grandmother_baseline = "Grandmother present at baseline", 
    grandmother_6m = "Grandmother present at 6 months", 
    grandmother_1y = "Grandmother present at 1 year",  
    famstruct_baseline = "Nuclear family structure (baseline)", 
    maritalscale_1y = "Marital quality scale at 1 year", 
    nonviolent_1y = "Husband non-violent at 1 year",  
    relationship_husb_1y = "Relationship with husband at 1 year",  
    relationship_inlaw_1y = "Relationship with mother in-law at 1 year", 
    no_kids_baseline = "Number of children at baseline", 
    c_first_child_baseline = "Index child is the first one", 
    girl_baseline = "Index child is female", 
    femshare_baseline = "Proportion of children ", 
    exclusivebf_6m = "Exclusive breastfeeding at 6 months", 
    anybf_6m = "Any breastfeeding at 6 months", 
    not_stunted_7y = "Not stunted at 7 years", 
    WAZ_6 = "Weight-for-age z-score at 6 months", 
    WAZ_1y = "Weight-for-age z-score at 1 year", 
    HAZ_6 = "Height-for-age z-score at 6 months", 
    HAZ_1y = "Height-for-age z-score at 1 year", 
    play_mo_1y = "Mother play frequency with infant at 1 year", 
    play_fa_1y = "Father play frequency with infant at 1 year"
  )



dat_trial_clean <- dat_trial_clean %>%
  select(c("response_6mo", 
     "response_1y", 
     "remission_6mo",
    "remission_1y", 
    "recovery_1y",
    "recovery_7y",
    "recover_never",
    "relapse_1y",
    "recurrence_7y",
    "treat",
    "attrit",
    "attrit2",
    "newid",
    "uc",
    "interviewer",
    "month_int",
    "depressed_6m", 
    "depressed_1y",
    "depressed_7y",
    "employed_fa_baseline", 
    "occupation_fa_baseline", 
    "electricity_baseline", 
    "radio_baseline", 
    "tv_baseline", 
    "fridge_baseline", 
    "bicycle_baseline", 
    "motorcycle_baseline", 
    "aircon_baseline", 
    "waterpump_baseline", 
    "cartruck_baseline", 
    "pipedwater_baseline", 
    "wellwithpump_baseline", 
    "wellwater_baseline", 
    "flushtoilet_baseline", 
    "anylatrine_baseline", 
    "brickwalls_baseline", 
    "enoughfoodmoney_baseline", 
    "metalroof_baseline",
    "basicneeds_baseline", 
    "enoughfood_baseline",  
    "employed_mo_baseline",
    "grandmother_baseline", 
    "grandmother_6m", 
    "grandmother_1y",  
    "famstruct_baseline",
    "nonviolent_1y", 
    "c_first_child_baseline", 
    "girl_baseline",
    "exclusivebf_6m", 
    "anybf_6m",
    "hamd_6m_percentchange_6m", 
    "hamd_1y_percentchange_1y", 
    "hamd_1y_percentchange_prev_1y", 
    "depindex_0", 
    "depindex_0_f", 
    "depindex_6m", 
    "depindex_1y", 
    "depindex_6m_f",
    "depindex_1y_f", 
    "age_baseline",
    "mo_wt_6m",
    "mo_ht_baseline", 
    "mo_bmi_baseline", 
    "hamd_baseline", 
    "hamd_baseline2",
    "hamd_6m", 
    "hamd_6m2",
    "hamd_1y", 
    "hamd_1y2",
    "bdq_baseline", 
    "bdq_6m", 
    "bdq_1y", 
    "gaf_baseline", 
    "gaf_6m", 
    "gaf_1y",  
    "ses_baseline", 
    "ses_6m", 
    "ses_1y", 
    "edu_mo_baseline", 
    "edu_fa_baseline",
    "wealth_baseline", 
    "ln_income_fa_6m", 
    "ln_income_fa_1y",
    "motherfinancial_baseline",
    "empowered_6m", 
    "mspss_baseline", 
    "mspss_6m", 
    "mspss_1y", 
    "maritalscale_1y",
    "relationship_husb_1y",  
    "relationship_inlaw_1y", 
    "no_kids_baseline",
    "femshare_baseline",
    "WAZ_6", 
    "WAZ_1y", 
    "HAZ_6", 
    "HAZ_1y", 
    "play_mo_1y", 
    "play_fa_1y"))

```


## Table 1 
```{r}

# Define variables to include for each timepoint
include_vars_6m <- c("treat", "hamd_baseline", "hamd_baseline2", "hamd_6m", "hamd_6m2", "hamd_6m_percentchange_6m", "bdq_baseline", "bdq_6m", "gaf_baseline", "gaf_6m", "mo_ht_baseline", "mo_bmi_baseline", "ses_baseline", "ses_6m",
                     "age_baseline", "edu_mo_baseline", "edu_fa_baseline", 
                     "employed_fa_baseline", "occupation_fa_baseline", "electricity_baseline", 
                     "radio_baseline", "tv_baseline", "fridge_baseline", "bicycle_baseline", 
                     "motorcycle_baseline", "aircon_baseline", "waterpump_baseline", 
                     "cartruck_baseline", "pipedwater_baseline", "wellwithpump_baseline", 
                     "wellwater_baseline", "flushtoilet_baseline", "anylatrine_baseline", 
                     "brickwalls_baseline", "enoughfoodmoney_baseline", "metalroof_baseline", 
                     "wealth_baseline", "ln_income_fa_6m", "basicneeds_baseline", 
                     "enoughfood_baseline", "employed_mo_baseline", "empowered_6m", 
                     "motherfinancial_baseline", "mspss_baseline", "mspss_6m", 
                     "grandmother_baseline", "grandmother_6m", "famstruct_baseline", "no_kids_baseline", 
                     "c_first_child_baseline", "girl_baseline", "femshare_baseline", "exclusivebf_6m", "anybf_6m", "WAZ_6", "HAZ_6")

include_vars_1y <- c("treat", "hamd_baseline", "hamd_baseline2",
                     "hamd_6m", "hamd_6m2", "hamd_6m_percentchange_6m", "hamd_1y", "hamd_1y2", "hamd_1y_percentchange_1y", "hamd_1y_percentchange_prev_1y", "bdq_baseline", "bdq_6m", 
                     "bdq_1y", "gaf_baseline", "gaf_6m", "gaf_1y", "mo_ht_baseline", "mo_wt_6m", "mo_bmi_baseline", "ses_baseline", 
                     "ses_6m", "ses_1y",  "age_baseline", "edu_mo_baseline", 
                     "edu_fa_baseline", "employed_fa_baseline", "occupation_fa_baseline", 
                     "electricity_baseline", "radio_baseline", "tv_baseline", 
                     "fridge_baseline", "bicycle_baseline", "motorcycle_baseline", 
                     "aircon_baseline", "waterpump_baseline", "cartruck_baseline", 
                     "pipedwater_baseline", "wellwithpump_baseline", "wellwater_baseline", 
                     "flushtoilet_baseline", "anylatrine_baseline", "brickwalls_baseline", 
                     "enoughfoodmoney_baseline", "metalroof_baseline", 
                     "wealth_baseline", "ln_income_fa_6m", "ln_income_fa_1y", "basicneeds_baseline", 
                     "enoughfood_baseline", "employed_mo_baseline", "empowered_6m", 
                     "motherfinancial_baseline", "mspss_baseline", "mspss_6m", "mspss_1y", 
                     "grandmother_baseline", "grandmother_6m", "grandmother_1y", "famstruct_baseline", "maritalscale_1y", "nonviolent_1y", 
                     "relationship_husb_1y", "relationship_inlaw_1y",
                     "no_kids_baseline", "c_first_child_baseline", "girl_baseline", 
                     "femshare_baseline", "exclusivebf_6m", "anybf_6m", "WAZ_6", "WAZ_1y", 
                     "HAZ_6", "HAZ_1y", "play_mo_1y", "play_fa_1y")

include_vars_7y <- c("treat", "hamd_baseline",  "hamd_baseline2", "hamd_6m", "hamd_6m2", "hamd_6m_percentchange_6m", "hamd_1y", "hamd_1y2", "hamd_1y_percentchange_1y", "hamd_1y_percentchange_prev_1y", "bdq_baseline", "bdq_6m", 
                     "bdq_1y", "gaf_baseline", "gaf_6m", "gaf_1y", "mo_ht_baseline", "mo_bmi_baseline", "mo_wt_6m", "ses_baseline", 
                     "ses_6m", "ses_1y", "age_baseline", "edu_mo_baseline", "edu_fa_baseline", 
                     "employed_fa_baseline", "occupation_fa_baseline", "electricity_baseline", 
                     "radio_baseline", "tv_baseline", "fridge_baseline", "bicycle_baseline", 
                     "motorcycle_baseline", "aircon_baseline", "waterpump_baseline", 
                     "cartruck_baseline", "pipedwater_baseline", "wellwithpump_baseline", 
                     "wellwater_baseline", "flushtoilet_baseline", "anylatrine_baseline", 
                     "brickwalls_baseline", "enoughfoodmoney_baseline", "metalroof_baseline", 
                     "wealth_baseline", "ln_income_fa_6m", 
                     "ln_income_fa_1y", "basicneeds_baseline", "enoughfood_baseline", 
                     "employed_mo_baseline", "empowered_6m", 
                     "motherfinancial_baseline", "mspss_baseline", "mspss_6m", 
                     "mspss_1y", "grandmother_baseline", "grandmother_6m", "grandmother_1y", 
                     "famstruct_baseline", "maritalscale_1y", "nonviolent_1y", 
                     "relationship_husb_1y", "relationship_inlaw_1y", "no_kids_baseline", 
                     "c_first_child_baseline", "girl_baseline", "femshare_baseline", 
                     "exclusivebf_6m", "anybf_6m", "WAZ_6", "WAZ_1y", 
                     "HAZ_6", "HAZ_1y", "play_mo_1y", "play_fa_1y")



# Create tbl_summary objects (one overall summary per timepoint)
tbl_6m <- dat_trial_clean %>% 
  #filter(attrit == 0 | attrit == 1) %>% ## Filter to 6 months
  filter(!is.na(remission_6mo)) %>%
  tbl_summary(
    by = remission_6mo,
    include = include_vars_6m,
    type = list(hamd_6m ~ "continuous",
                hamd_6m2 ~ "continuous",
                hamd_6m_percentchange_6m ~ "continuous",
                no_kids_baseline ~ "continuous",
                HAZ_6 ~ "continuous"),
    missing = "ifany",
    percent = "column",
    statistic = list(all_continuous() ~ "{mean} ({sd}) | {median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      all_categorical() ~ c(0, 0),
      all_continuous() ~ c(1, 1)
  )
  ) %>%
  add_p(pvalue_fun  = function(x) style_pvalue(x, digits = 3)) %>%
  bold_p(t = 0.05) %>%
  bold_labels() 


tbl_1y <- dat_trial_clean %>% 
  #filter(attrit == 0 | attrit == 1) %>% # Filter to 1 year
  filter(!is.na(relapse_1y)) %>%
  tbl_summary(
    by = relapse_1y,
    include = include_vars_1y,
    type = list(hamd_6m ~ "continuous",
                hamd_6m2 ~ "continuous",
                hamd_6m_percentchange_6m ~ "continuous",
                no_kids_baseline ~ "continuous",
                bdq_1y ~ "continuous",
                HAZ_6 ~ "continuous",
                HAZ_1y ~ "continuous",
                hamd_1y_percentchange_1y ~ "continuous",
                hamd_1y_percentchange_prev_1y ~ "continuous",
                relationship_husb_1y ~ "continuous",
                relationship_inlaw_1y ~ "continuous",
                play_mo_1y ~ "continuous",
                play_fa_1y ~ "continuous"),
    missing = "ifany",
    percent = "column",
    statistic = list(all_continuous() ~ "{mean} ({sd}) | {median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      all_categorical() ~ c(0, 0),
      all_continuous() ~ c(1, 1)
  )
  ) %>%
  add_p(pvalue_fun  = function(x) style_pvalue(x, digits = 3)) %>%
  bold_p(t = 0.05) %>%
  bold_labels() 


tbl_7y <- dat_trial_clean %>% 
  #filter(attrit2 == 0 & attrit == 0) %>% # Filter to 7 years
  filter(!is.na(recurrence_7y)) %>%
  tbl_summary(
    by = recurrence_7y,
    include = include_vars_7y,
    type = list(hamd_6m ~ "continuous",
                hamd_6m2 ~ "continuous",
                hamd_6m_percentchange_6m ~ "continuous",
                no_kids_baseline ~ "continuous",
                bdq_1y ~ "continuous",
                HAZ_6 ~ "continuous",
                HAZ_1y ~ "continuous",
                hamd_1y_percentchange_1y ~ "continuous",
                hamd_1y_percentchange_prev_1y ~ "continuous",
                relationship_husb_1y ~ "continuous",
                relationship_inlaw_1y ~ "continuous",
                play_mo_1y ~ "continuous",
                play_fa_1y ~ "continuous"),
    missing = "ifany",
    percent = "column",
    statistic = list(all_continuous() ~ "{mean} ({sd}) | {median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(
      all_categorical() ~ c(0, 0),
      all_continuous() ~ c(1, 1)
  )
  ) %>%
  add_p(pvalue_fun  = function(x) style_pvalue(x, digits = 3)) %>%
  bold_p(t = 0.05) %>%
  bold_labels() 

# Merge the summary tables side-by-side
merged_tbl <- tbl_merge(
  list(tbl_6m, tbl_1y, tbl_7y),
  tab_spanner = c("Remission at 6 months", "Relapse at 1 year", "Recurrence at 7 years")
)

merged_tbl
# Convert to flextable for further formatting and export
merged_tbl_flex <- as_flex_table(merged_tbl) %>%
  theme_vanilla() %>%
  autofit() %>%
  fontsize(size = 10, part = "all") %>%
  set_caption("Merged Summary Table by Timepoint (Horizontal Format)")

# Save to a Word document
save_as_docx(merged_tbl_flex, path = "outputs/merged_summary_horizontal.docx")

# Optionally, view the flextable in RStudio
merged_tbl_flex

```


# ─────────────────────────────────────────────
# Section 3: Subset data for ML
# ─────────────────────────────────────────────

### Relapse at one year
```{r}

#### Remove predictors assessed at one and seven years
dat_ml_1y <- dat_trial_clean %>%
  #filter(attrit == 0 | attrit == 1) %>% # Filter to 1 year
  filter(!is.na(relapse_1y)) %>% #### Drop NA outcome
  dplyr::select(c(treat, newid, uc, interviewer, month_int,
remission_6mo, hamd_6m_percentchange_6m, response_6mo, relapse_1y, depindex_0, depindex_6m, depindex_6m_f, depressed_6m, depindex_0, depindex_0_f, depindex_6m, mo_ht_baseline, mo_wt_6m, mo_bmi_baseline, hamd_baseline, hamd_6m, bdq_baseline, bdq_6m, gaf_baseline, gaf_6m,  ses_baseline, ses_6m, age_baseline, edu_mo_baseline, edu_fa_baseline, employed_fa_baseline, occupation_fa_baseline, electricity_baseline, radio_baseline, tv_baseline, fridge_baseline, bicycle_baseline, motorcycle_baseline, aircon_baseline, waterpump_baseline, cartruck_baseline, pipedwater_baseline, wellwithpump_baseline, wellwater_baseline, flushtoilet_baseline, anylatrine_baseline, brickwalls_baseline, enoughfoodmoney_baseline, metalroof_baseline, wealth_baseline, ln_income_fa_6m, basicneeds_baseline, enoughfood_baseline, employed_mo_baseline,  empowered_6m, motherfinancial_baseline, mspss_baseline, mspss_6m, grandmother_baseline, grandmother_6m, famstruct_baseline,  no_kids_baseline, c_first_child_baseline, girl_baseline, femshare_baseline, exclusivebf_6m, anybf_6m, WAZ_6, HAZ_6))

# Identify variables with haven labels
problematic_vars <- sapply(dat_ml_1y, is.labelled)
print(names(dat_ml_1y)[problematic_vars])

# Function to convert selected variables to numeric
convert_to_numeric <- function(data, var_list) {
  data <- data %>% mutate(across(all_of(var_list), ~ as.numeric(as.character(haven::zap_labels(.)))))
  return(data)
}

# List of variables to convert
#numeric_vars <- c("grandmother_1y")

# Apply conversion
#dat_ml_1y <- convert_to_numeric(dat_ml_1y, numeric_vars)
#str(dat_ml_1y[numeric_vars])


# Convert matrix cols to vector
matrix_columns <- names(dat_ml_1y)[sapply(dat_ml_1y, is.matrix)]
print(matrix_columns)

dat_ml_1y <- dat_ml_1y %>%
  mutate(across(all_of(matrix_columns), ~ as.numeric(.)))

## Export non-imputed dataset
write_csv(dat_ml_1y, "dat_ml_1y.csv")
dat_ml_1y <- read_csv("dat_ml_1y.csv")

```


### Recurrence at 7 years
```{r}

dat_ml_7y <- dat_trial_clean %>%
  #filter(attrit2 == 0 & attrit == 0) %>% # Filter to 7 years
  filter(!is.na(recurrence_7y)) %>% #### Drop NA outcome
  dplyr::select(c(treat, newid, uc, interviewer, month_int, recurrence_7y, remission_1y, 
remission_6mo, hamd_6m_percentchange_6m, hamd_1y_percentchange_1y, hamd_1y_percentchange_prev_1y, response_6mo, response_1y, recovery_1y, depindex_0, depindex_6m, depindex_1y, depindex_6m_f, depressed_6m, depressed_1y, depindex_0, depindex_0_f, depindex_6m, depindex_1y_f, mo_ht_baseline, mo_bmi_baseline, mo_wt_6m, hamd_baseline, hamd_6m, hamd_1y, bdq_baseline, bdq_6m, bdq_1y, gaf_baseline, gaf_6m, gaf_1y,  ses_baseline, ses_6m, ses_1y, edu_mo_baseline, edu_fa_baseline, employed_fa_baseline, occupation_fa_baseline, electricity_baseline, radio_baseline, tv_baseline, fridge_baseline, bicycle_baseline, motorcycle_baseline, aircon_baseline, waterpump_baseline, cartruck_baseline, pipedwater_baseline, wellwithpump_baseline, wellwater_baseline, flushtoilet_baseline, anylatrine_baseline, brickwalls_baseline, enoughfoodmoney_baseline, metalroof_baseline, wealth_baseline, ln_income_fa_6m, ln_income_fa_1y, basicneeds_baseline, enoughfood_baseline,  employed_mo_baseline, empowered_6m, mspss_baseline, mspss_6m, mspss_1y,  grandmother_baseline, grandmother_6m, grandmother_1y,  famstruct_baseline, maritalscale_1y, nonviolent_1y,  relationship_husb_1y,  relationship_inlaw_1y, no_kids_baseline, c_first_child_baseline, girl_baseline, femshare_baseline, exclusivebf_6m, anybf_6m, WAZ_6, WAZ_1y, HAZ_6, HAZ_1y, play_mo_1y, play_fa_1y))

# Identify variables with haven labels
problematic_vars <- sapply(dat_ml_7y, is.labelled)
print(names(dat_ml_7y)[problematic_vars])

# Function to convert selected variables to numeric
convert_to_numeric <- function(data, var_list) {
  data <- data %>% mutate(across(all_of(var_list), ~ as.numeric(as.character(haven::zap_labels(.)))))
  return(data)
}

# List of variables to convert
#numeric_vars <- c("grandmother_7y", "genhealth_7y", "income_hh_7y", "income_mo_7y")

# Apply conversion
#dat_ml_7y <- convert_to_numeric(dat_ml_7y, numeric_vars)
#str(dat_ml_7y[numeric_vars])


# Convert matrix cols to vector
matrix_columns <- names(dat_ml_7y)[sapply(dat_ml_7y, is.matrix)]
print(matrix_columns)

dat_ml_7y <- dat_ml_7y %>%
  mutate(across(all_of(matrix_columns), ~ as.numeric(.)))

## Export non-imputed dataset
write_csv(dat_ml_7y, "dat_ml_7y.csv")
dat_ml_7y <- read_csv("dat_ml_7y.csv")


```



### Remission at 6 months
```{r}

#### Remove predictors assessed at 6 months, 1 year, and  years
dat_ml_6m <- dat_trial_clean %>%
  #filter(attrit == 0 | attrit == 1) %>% # Filter to 6 months
  filter(!is.na(remission_6mo)) %>% #### Drop NA outcome
  dplyr::select(c(treat, newid, uc, interviewer, month_int, remission_6mo, depindex_0, depindex_0_f, mo_ht_baseline, mo_bmi_baseline, hamd_baseline, bdq_baseline, gaf_baseline, ses_baseline, age_baseline, edu_mo_baseline, edu_fa_baseline, employed_fa_baseline, occupation_fa_baseline, electricity_baseline, radio_baseline, tv_baseline, fridge_baseline, bicycle_baseline, motorcycle_baseline, aircon_baseline, waterpump_baseline, cartruck_baseline, pipedwater_baseline, wellwithpump_baseline, wellwater_baseline, flushtoilet_baseline, anylatrine_baseline, brickwalls_baseline, enoughfoodmoney_baseline, metalroof_baseline, wealth_baseline, basicneeds_baseline, enoughfood_baseline, employed_mo_baseline, mspss_baseline, grandmother_baseline, famstruct_baseline, no_kids_baseline, c_first_child_baseline, girl_baseline, femshare_baseline)) #%>%
  #drop_na()  

# Identify variables with haven labels
problematic_vars <- sapply(dat_ml_6m, is.labelled)
print(names(dat_ml_6m)[problematic_vars])

# Function to convert selected variables to numeric
convert_to_numeric <- function(data, var_list) {
  data <- data %>% mutate(across(all_of(var_list), ~ as.numeric(as.character(haven::zap_labels(.)))))
  return(data)
}

# List of variables to convert
#numeric_vars <- c("grandmother_6m")

# Apply conversion
#dat_ml_6m <- convert_to_numeric(dat_ml_6m, numeric_vars)
#str(dat_ml_6m[numeric_vars])


# Convert matrix cols to vector
matrix_columns <- names(dat_ml_6m)[sapply(dat_ml_6m, is.matrix)]
print(matrix_columns)

dat_ml_6m <- dat_ml_6m %>%
  mutate(across(all_of(matrix_columns), ~ as.numeric(.)))


## Export non-imputed dataset
write_csv(dat_ml_6m, "dat_ml_6m.csv")
dat_ml_6m <- read_csv("dat_ml_6m.csv")

```

### Depressed at 7 years
```{r}

dat_ml_7y_dep <- dat_trial_clean %>%
  #filter(attrit2 == 0 & attrit == 0) %>% # Filter to 7 years
  filter(!is.na(depressed_7y)) %>% #### Drop NA outcome
  dplyr::select(c(treat, newid, uc, interviewer, month_int, depressed_7y, recurrence_7y, remission_1y, 
remission_6mo, hamd_6m_percentchange_6m, hamd_1y_percentchange_1y, hamd_1y_percentchange_prev_1y, response_6mo, response_1y, recovery_1y, depindex_0, depindex_6m, depindex_1y, depindex_6m_f, depressed_6m, depressed_1y, depindex_0, depindex_0_f, depindex_6m, depindex_1y_f, mo_ht_baseline, mo_bmi_baseline, mo_wt_6m, hamd_baseline, hamd_6m, hamd_1y, bdq_baseline, bdq_6m, bdq_1y, gaf_baseline, gaf_6m, gaf_1y,  ses_baseline, ses_6m, ses_1y, edu_mo_baseline, edu_fa_baseline, employed_fa_baseline, occupation_fa_baseline, electricity_baseline, radio_baseline, tv_baseline, fridge_baseline, bicycle_baseline, motorcycle_baseline, aircon_baseline, waterpump_baseline, cartruck_baseline, pipedwater_baseline, wellwithpump_baseline, wellwater_baseline, flushtoilet_baseline, anylatrine_baseline, brickwalls_baseline, enoughfoodmoney_baseline, metalroof_baseline, wealth_baseline, ln_income_fa_6m, ln_income_fa_1y, basicneeds_baseline, enoughfood_baseline,  employed_mo_baseline, empowered_6m, mspss_baseline, mspss_6m, mspss_1y,  grandmother_baseline, grandmother_6m, grandmother_1y,  famstruct_baseline, maritalscale_1y, nonviolent_1y,  relationship_husb_1y,  relationship_inlaw_1y, no_kids_baseline, c_first_child_baseline, girl_baseline, femshare_baseline, exclusivebf_6m, anybf_6m, WAZ_6, WAZ_1y, HAZ_6, HAZ_1y, play_mo_1y, play_fa_1y)) 

# Identify variables with haven labels
problematic_vars <- sapply(dat_ml_7y_dep, is.labelled)
print(names(dat_ml_7y_dep)[problematic_vars])

# Function to convert selected variables to numeric
convert_to_numeric <- function(data, var_list) {
  data <- data %>% mutate(across(all_of(var_list), ~ as.numeric(as.character(haven::zap_labels(.)))))
  return(data)
}

# List of variables to convert
#numeric_vars <- c("grandmother_7y", "genhealth_7y", "income_hh_7y", "income_mo_7y")

# Apply conversion
#dat_ml_7y_dep <- convert_to_numeric(dat_ml_7y_dep, numeric_vars)
#str(dat_ml_7y_dep[numeric_vars])


# Convert matrix cols to vector
matrix_columns <- names(dat_ml_7y_dep)[sapply(dat_ml_7y_dep, is.matrix)]
print(matrix_columns)

dat_ml_7y_dep <- dat_ml_7y_dep %>%
  mutate(across(all_of(matrix_columns), ~ as.numeric(.)))

## Export non-imputed dataset
write_csv(dat_ml_7y_dep, "dat_ml_7y_dep.csv")
dat_ml_7y_dep <- read_csv("dat_ml_7y_dep.csv")

```

### Never recovered at 7 years
```{r}

dat_ml_7y_never <- dat_trial_clean %>%
  #filter(attrit2 == 0 & attrit == 0) %>% # Filter to 7 years
  filter(!is.na(recover_never)) %>% #### Drop NA outcome
  dplyr::select(c(treat, newid, uc, interviewer, month_int, recover_never, recurrence_7y, remission_1y, 
remission_6mo, hamd_6m_percentchange_6m, hamd_1y_percentchange_1y, hamd_1y_percentchange_prev_1y, response_6mo, response_1y, recovery_1y, depindex_0, depindex_6m, depindex_1y, depindex_6m_f, depressed_6m, depressed_1y, depindex_0, depindex_0_f, depindex_6m, depindex_1y_f, mo_ht_baseline, mo_bmi_baseline, mo_wt_6m, hamd_baseline, hamd_6m, hamd_1y, bdq_baseline, bdq_6m, bdq_1y, gaf_baseline, gaf_6m, gaf_1y,  ses_baseline, ses_6m, ses_1y, edu_mo_baseline, edu_fa_baseline, employed_fa_baseline, occupation_fa_baseline, electricity_baseline, radio_baseline, tv_baseline, fridge_baseline, bicycle_baseline, motorcycle_baseline, aircon_baseline, waterpump_baseline, cartruck_baseline, pipedwater_baseline, wellwithpump_baseline, wellwater_baseline, flushtoilet_baseline, anylatrine_baseline, brickwalls_baseline, enoughfoodmoney_baseline, metalroof_baseline, wealth_baseline, ln_income_fa_6m, ln_income_fa_1y, basicneeds_baseline, enoughfood_baseline,  employed_mo_baseline, empowered_6m, mspss_baseline, mspss_6m, mspss_1y,  grandmother_baseline, grandmother_6m, grandmother_1y,  famstruct_baseline, maritalscale_1y, nonviolent_1y,  relationship_husb_1y,  relationship_inlaw_1y, no_kids_baseline, c_first_child_baseline, girl_baseline, femshare_baseline, exclusivebf_6m, anybf_6m, WAZ_6, WAZ_1y, HAZ_6, HAZ_1y, play_mo_1y, play_fa_1y)) 

# Identify variables with haven labels
problematic_vars <- sapply(dat_ml_7y_never, is.labelled)
print(names(dat_ml_7y_never)[problematic_vars])

# Function to convert selected variables to numeric
convert_to_numeric <- function(data, var_list) {
  data <- data %>% mutate(across(all_of(var_list), ~ as.numeric(as.character(haven::zap_labels(.)))))
  return(data)
}

# List of variables to convert
#numeric_vars <- c("grandmother_7y", "genhealth_7y", "income_hh_7y", "income_mo_7y")

# Apply conversion
#dat_ml_7y_never <- convert_to_numeric(dat_ml_7y_never, numeric_vars)
#str(dat_ml_7y_never[numeric_vars])


# Convert matrix cols to vector
matrix_columns <- names(dat_ml_7y_never)[sapply(dat_ml_7y_never, is.matrix)]
print(matrix_columns)

dat_ml_7y_never <- dat_ml_7y_never %>%
  mutate(across(all_of(matrix_columns), ~ as.numeric(.)))

## Export non-imputed dataset
write_csv(dat_ml_7y_never, "dat_ml_7y_never.csv")
dat_ml_7y_never <- read_csv("dat_ml_7y_never.csv")

```


## Diagram of depression trajectories
```{r}

# install.packages(c("ggplot2","grid"))
library(grid)

# ── 1) Main “successful” trajectory (solid) ────────────────────────────────
main <- data.frame(
  time     = c(0,  3,   6,   36,  84),
  severity = c(20, 10,   5,    5,   5)
)

# 2) Landmark events on the main path
landmarks <- data.frame(
  time     = c( 3,   6,   36),
  severity = c(10,   5,    5),
  label    = c("Response", "Remission", "Recovery"),
  shape    = 4
)

# 3) Relapse arc (dashed) from Remission → Relapse
relapse_arc <- data.frame(
  x    = 6,    y    = 5,
  xend = 12,   yend = 18
)
relapse_pt  <- data.frame(time=12, severity=18, label="Relapse", shape=1)

# 4) Recurrence arc (dashed) from Recovery → Recurrence
recurr_arc <- data.frame(
  x    = 36,   y    = 5,
  xend = 84,   yend = 17
)
recurr_pt  <- data.frame(time=84, severity=17, label="Recurrence", shape=1)

# ── 5) Plot ────────────────────────────────────────────────────────────────
# Compute the vertical limits (add a little padding)
# before plotting
y_min <- min(main$severity) * 0.95
y_max <- max(main$severity) * 1.05

ggplot() +
  # 1) custom panel box
  annotate("rect",
           xmin = 0,   xmax = 90,
           ymin = y_min, ymax = y_max,
           fill  = NA,
           color = "black",
           size  = 1) +
  
  # 2) main trajectory
  geom_line(data = main,
            aes(time, severity),
            size  = 1.5,
            color = "#2c7fb8") +
  
  # 3) event markers
  geom_point(data = landmarks,
             aes(time, severity),
             shape  = 4,
             size   = 4,
             stroke = 1.2) +
  geom_text(data = landmarks,
            aes(time, severity, label = label),
            vjust = -1.2,
            size  = 3) +
  
  # 4) relapse arc
  geom_curve(data = relapse_arc,
             aes(x = x, y = y, xend = xend, yend = yend),
             curvature = 0.3,
             linetype  = "dashed",
             color     = "gray40") +
  geom_point(data = relapse_pt,
             aes(time, severity),
             shape  = 1,
             size   = 4,
             stroke = 1.2) +
  geom_text(data = relapse_pt,
            aes(time, severity, label = label),
            vjust = -1.2,
            size  = 3) +
  
  # 5) recurrence arc
  geom_curve(data = recurr_arc,
             aes(x = x, y = y, xend = xend, yend = yend),
             curvature = 0.3,
             linetype  = "dashed",
             color     = "gray40") +
  geom_point(data = recurr_pt,
             aes(time, severity),
             shape  = 1,
             size   = 4,
             stroke = 1.2) +
  geom_text(data = recurr_pt,
            aes(time, severity, label = label),
            vjust = -1.2,
            size  = 3) +
  
  # 6) MDD threshold line + label
  geom_hline(yintercept = 7,
             linetype    = "dashed",
             color       = "firebrick",
             size        = 0.8) +
  annotate("text",
           x     = 5,
           y     = 8,
           label = "Threshold for\nMDD diagnosis",
           color = "firebrick",
           size  = 3,
           hjust = 0) +
  
  # 7) axes
  scale_x_continuous(
    name   = "Time since baseline",
    breaks = c(6, 12, 84),
    labels = c("6 mo", "1 yr", "7 yr"),
    limits = c(0, 90),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    name   = "Depression severity\n(higher = more severe)",
    breaks = NULL,
    expand = expansion(mult = c(0, 0.05))
  ) +
  
  coord_cartesian(clip = "off") +
  
  # 8) theme
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major   = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.title.x       = element_blank(),
    axis.title.y       = element_text(face = "bold", margin = margin(r = 10)),
    axis.text.x        = element_text(color = "black"),
    axis.ticks.length  = unit(3, "pt"),
    legend.position    = "right",
    legend.title       = element_blank(),
    legend.text        = element_text(size = 12),
    plot.margin        = margin(10, 10, 10, 10)
  )

```

